## Exchangable Magnesium Map Modelling and Testing

-   selected mods: MLR, OK, MLRK, SVM

### Data Preparation
#### set project directory and projections

```{r, echo=FALSE}
proj_dir <- 'D:/r/KenyaDSM'
wsg84 <- "+init=epsg:4326"
utm <- "+init=epsg:32236"
library(caret)
library(raster)
library(e1071)
library(sp)
library(RColorBrewer)
library(gstat)
library(automap)
```

#### select potassium data and training covariates

```{r}
soil_path <- paste(proj_dir,'data/soil/ke_sp/sample', sep="/")
mg_dat <- as.data.frame(read.csv(paste(soil_path, "mg_model.csv", sep = "/"), header = TRUE, sep = ","))
covs_path <- paste(proj_dir,'data/covs/local/', sep="/")
mg_covs <-  names(mg_dat[1:9])
mg_files <- paste0(covs_path,"/",mg_covs, ".tif")
mg_covs <- stack(mg_files)
summary(mg_covs)
plot(mg_covs$land_elevation)
```

#### log transformation

```{r}
hist(mg_dat$ExMg) # right skewed
hist(log(mg_dat$ExMg)) #log transformation of targets for normal distribution
mg_dat$log_ExMg <- log(mg_dat$ExMg)
```

#### split into 80:20 training/testing

```{r}
part <- createDataPartition(mg_dat$log_ExMg, times = 1, p = 0.2, list = FALSE)
mg_test <- mg_dat[part, ]
mg_train <- mg_dat[-part, ]
head(mg_train)
head(mg_test)
```

#### set equations

```{r}
mg_regeqn <- as.formula(log_ExMg ~ annual_ppt+bareground+driest_ppt+land_elevation+terrain_slope+treecover+valley_depth+wetness_index+wettest_ppt)
mg_okeqn <- as.formula(log_ExMg ~ 1)
```

### MLR Modelling

#### run model
```{r}
mg_mlr <- train(mg_regeqn, data = mg_train, method= "lm")
mg_mlr
```

### SVM Regression

#### run initial model

```{r}
mg_svm <- train(mg_regeqn, data = mg_train, method = 'svmRadial')
mg_svm
```

#### tune hyperparams
```{r}
svm_tune <- tune(svm, 
                 mg_regeqn, 
                 data = mg_train[,c("log_ExMg",names(mg_covs))],
                 ranges = list(
                   gamma = seq(0.2,0.5,0.05),
                   cost = seq(0.1,1,0.1)))
plot(svm_tune) # towards 0.5 gamma, 0.4 cost
svm_tune$best.parameters
```

#### remodel with best params

```{r}
hp <- expand.grid(C = svm_tune$best.parameters$cost,sigma = svm_tune$best.parameters$gamma)
mg_svm <- train(mg_regeqn, data = mg_train, method = 'svmRadial', tuneGrid = hp)
mg_svm
```
#### generate MLR and SVM maps
```{r}
mg_mlrpred <- exp(predict(mg_covs, mg_mlr))# MLR
mg_svmpred <- exp(predict(mg_covs, mg_svm))# SVM
```

### OK

#### convert to spatial format for kriging: change projection to UTM

```{r}
coordinates(mg_train) <- ~ X + Y
mg_train@proj4string <- CRS(wsg84)
mg_train <- spTransform(mg_train, CRS(utm))
mg_covs <- projectRaster(mg_covs, crs = CRS(utm))
mg_covs <- as(mg_covs, "SpatialGridDataFrame")
mg_train <- mg_train[which(!duplicated(mg_train@coords)),] #remove any duplicates
```

#### run autokrige

```{r}
mg_ok <- autoKrige(mg_okeqn, 
                  input_data = mg_train, 
                  verbose = TRUE)
plot(mg_ok)
```

### MLRK

#### run autokrige
```{r}
mg_mlrk <- autoKrige(mg_regeqn, 
                  input_data = mg_train, 
                  new_data = mg_covs,
                  verbose = TRUE)
plot(mg_mlrk)
```

### Generate predicted kriging maps

```{r}
mg_okpred <- exp(raster(mg_ok$krige_output[1])) #OK
mg_mlrkpred <- exp(raster(mg_mlrk$krige_output[1])) #MLRK

# convert projection to WGS84
mg_mlrpred <- projectRaster(mg_mlrpred, crs = wsg84)
mg_svmpred <- projectRaster(mg_svmpred, crs = wsg84)
mg_okpred <- projectRaster(mg_okpred, crs = wsg84)
mg_mlrkpred <- projectRaster(mg_mlrkpred, crs = wsg84)

{
  plot(mg_mlrpred)
  title("Prediction by MLR")
}
{
  plot(mg_svmpred)
  title("Prediction by SVM")
}
{
  plot(mg_okpred)
  title("Prediction by OK")
}
{
  plot(mg_mlrkpred)
  title("Prediction by MLRK")
}
```
#### export models
```{r}
model_path <- paste(proj_dir,"models", sep="/")
saveRDS(mg_mlr, file = paste(model_path, "mg_mlr.Rds", sep="/"))
saveRDS(mg_svm, file = paste(model_path, "mg_svm.Rds", sep="/"))
saveRDS(mg_ok, file = paste(model_path, "mg_ok.Rds", sep="/"))
saveRDS(mg_mlrk, file = paste(model_path, "mg_mlrk.Rds", sep="/"))
```

#### export generated maps

```{r, echo=FALSE}
model_path <- paste(proj_dir,"data/soil/models", sep="/")
setwd(model_path)
writeRaster(mg_mlrpred, filename = "mgmlr_pred.tif", overwrite = T)
writeRaster(mg_svmpred, filename = "mgsvm_pred.tif", overwrite = T)
writeRaster(mg_okpred, filename = "mgok_pred.tif", overwrite = T)
writeRaster(mg_mlrkpred, filename = "mgmlrk_pred.tif", overwrite = T)
```

## Model Testing

### select test data and convert to sp; select model rasters

```{r}
coordinates(mg_test) <- ~ X + Y
mg_test@proj4string <- CRS(wsg84)

mg_svmpred <- raster(paste(model_path,"mgsvm_pred.tif", sep = "/"))
mg_mlrpred <- raster(paste(model_path,"mgmlr_pred.tif", sep = "/" ))
mg_okpred <- raster(paste(model_path,"mgok_pred.tif", sep = "/"))
mg_mlrkpred <- raster(paste(model_path,"mgmlrk_pred.tif", sep = "/"))
```

### extract predicted values from rasters

```{r}
mg_test <- extract(x = mg_svmpred, y = mg_test, sp = TRUE)
mg_test <- extract(x = mg_mlrpred, y = mg_test, sp = TRUE)
mg_test <- extract(x = mg_okpred, y = mg_test, sp = TRUE)
mg_test <- extract(x = mg_mlrkpred, y = mg_test, sp = TRUE)
```

### calculate predicted errors

```{r}
mg_test$PE_svm <- mg_test$mgsvm_pred - mg_test$ExMg
mg_test$PE_mlr <- mg_test$mgmlr_pred - mg_test$ExMg
mg_test$PE_ok <- mg_test$mgok_pred - mg_test$ExMg
mg_test$PE_mlrk <- mg_test$mgmlrk_pred - mg_test$ExMg
summary(mg_test)
```

### calculate RMSE
```{r}
RMSE_svm <- round(sqrt(sum(mg_test$PE_svm^2, na.rm=TRUE) / length(mg_test$PE_svm)),2)
RMSE_mlr <- round(sqrt(sum(mg_test$PE_mlr^2, na.rm=TRUE) / length(mg_test$PE_mlr)),2)
RMSE_ok <- round(sqrt(sum(mg_test$PE_ok^2, na.rm=TRUE) / length(mg_test$PE_ok)), 2)
RMSE_mlrk <- round(sqrt(sum(mg_test$PE_mlrk^2, na.rm=TRUE) / length(mg_test$PE_mlrk)),2)

paste0("svm: ", RMSE_svm)
paste0("mlr: ", RMSE_mlr)
paste0("mlrk: ", RMSE_mlrk)
paste0("ok: ", RMSE_ok)
```