## Exchangable Potassium Map Modelling and Testing

-   selected mods: MLR, OK, MLRK, SVM

### Data Preparation

#### set project directory and projections

```{r, echo=FALSE}
proj_dir <- 'D:/r/KenyaDSM'
wsg84 <- "+init=epsg:4326"
utm <- "+init=epsg:32236"
library(caret)
library(raster)
library(e1071)
library(sp)
library(RColorBrewer)
library(gstat)
library(automap)
```

#### select potassium data and training covariates

```{r}
soil_path <- paste(proj_dir,'data/soil/ke_sp/sample', sep="/")
k_dat <- as.data.frame(read.csv(paste(soil_path, "k_model.csv", sep = "/"), header = TRUE, sep = ","))
covs_path <- paste(proj_dir,'data/covs/local/', sep="/")
k_covs <-  names(k_dat[1:9])
k_files <- paste0(covs_path,"/",k_covs, ".tif")
k_covs <- stack(k_files)
summary(k_covs)
plot(k_covs$land_elevation)
```

#### log transformation

```{r}
hist(k_dat$ExK) # right skewed
hist(log(k_dat$ExK)) #log transformation of targets for normal distribution
k_dat$log_ExK <- log(k_dat$ExK)
```

#### split into 80:20 training/testing

```{r}
part <- createDataPartition(k_dat$log_ExK, times = 1, p = 0.2, list = FALSE)
k_test <- k_dat[part, ]
k_train <- k_dat[-part, ]
head(k_train)
head(k_test)
```

#### set equations

```{r}
k_regeqn <- as.formula(log_ExK ~ annual_ppt+bareground+land_elevation+treecover+valley_depth+veg_index+wetness_index+wettest_ppt)
k_okeqn <- as.formula(log_ExK ~ 1)
```

### MLR Modelling

#### run model

```{r}
k_mlr <- train(k_regeqn, data = k_train, method= "lm")
k_mlr
```

### SVM Regression

#### run initial model

```{r}
k_svm <- train(k_regeqn, data = k_train, method = 'svmRadial')
k_svm # RMSE - 1.21
```

#### tune hyperparams

```{r}
svm_tune <- tune(svm, 
                 k_regeqn, 
                 data = k_train[,c("log_ExK",names(k_covs))],
                 ranges = list(
                   gamma = seq(0.2,0.5,0.05),
                   cost = seq(0.1,1,0.1)))
plot(svm_tune) # towards 0.5 gamma, 0.3 cost
svm_tune$best.parameters
```

#### remodel with best params

```{r}
hp <- expand.grid(C = svm_tune$best.parameters$cost,sigma = svm_tune$best.parameters$gamma)
k_svm <- train(k_regeqn, data = k_train, method = 'svmRadial', tuneGrid = hp)
k_svm
```

#### generate MLR and SVM maps

```{r}
k_mlrpred <- exp(predict(k_covs, k_mlr))# MLR
k_svmpred <- exp(predict(k_covs, k_svm))# SVM
```

### OK

#### convert to spatial format for kriging: change projection to UTM

```{r}
coordinates(k_train) <- ~ X + Y
k_train@proj4string <- CRS(wsg84)
k_train <- spTransform(k_train, CRS(utm))
k_covs <- projectRaster(k_covs, crs = CRS(utm))
k_covs <- as(k_covs, "SpatialGridDataFrame")
k_train <- k_train[which(!duplicated(k_train@coords)),] #remove any duplicates
```

#### run autokrige

```{r}
k_ok <- autoKrige(k_okeqn, 
                  input_data = k_train, 
                  verbose = TRUE)
plot(k_ok)
```

### MLRK

#### run autokrige

```{r}
k_mlrk <- autoKrige(k_regeqn, 
                  input_data = k_train, 
                  new_data = k_covs,
                  verbose = TRUE)
plot(k_mlrk)
```

### Generate kriging predicted maps

```{r}
k_okpred <- exp(raster(k_ok$krige_output[1])) #OK
k_mlrkpred <- exp(raster(k_mlrk$krige_output[1])) #MLRK

# convert projection to WGS84
k_mlrpred <- projectRaster(k_mlrpred, crs = wsg84)
k_svmpred <- projectRaster(k_svmpred, crs = wsg84)
k_okpred <- projectRaster(k_okpred, crs = wsg84)
k_mlrkpred <- projectRaster(k_mlrkpred, crs = wsg84)

{
  plot(k_mlrpred)
  title("Prediction by MLR")
}
{
  plot(k_svmpred)
  title("Prediction by SVM")
}
{
  plot(k_okpred)
  title("Prediction by OK")
}
{
  plot(k_mlrkpred)
  title("Prediction by MLRK")
}
```

#### export models

```{r}
model_path <- paste(proj_dir,"maps/models", sep="/")
saveRDS(k_mlr, file = paste(model_path, "k_mlr.Rds", sep="/"))
saveRDS(k_svm, file = paste(model_path, "k_svm.Rds", sep="/"))
saveRDS(k_ok, file = paste(model_path, "k_ok.Rds", sep="/"))
saveRDS(k_mlrk, file = paste(model_path, "k_mlrk.Rds", sep="/"))
```

#### export generated maps

```{r, echo=FALSE}
model_path <- paste(proj_dir,"data/soil/models", sep="/")
setwd(model_path)
writeRaster(k_mlrpred, filename = "kmlr_pred.tif", overwrite=T)
writeRaster(k_svmpred, filename = "ksvm_pred.tif", overwrite=T)
writeRaster(k_okpred, filename = "kok_pred.tif", overwrite=T)
writeRaster(k_mlrkpred, filename = "kmlrk_pred.tif", overwrite=T)
```

## Model Testing

### select test data and convert to sp; select model rasters

```{r}
coordinates(k_test) <- ~ X + Y
k_test@proj4string <- CRS(wsg84)

k_svmpred <- raster(paste(model_path,"ksvm_pred.tif", sep = "/"))
k_mlrpred <- raster(paste(model_path,"kmlr_pred.tif", sep = "/" ))
k_okpred <- raster(paste(model_path,"kok_pred.tif", sep = "/"))
k_mlrkpred <- raster(paste(model_path,"kmlrk_pred.tif", sep = "/"))
```

### extract predicted values from rasters

```{r}
k_test <- extract(x = k_svmpred, y = k_test, sp = TRUE)
k_test <- extract(x = k_mlrpred, y = k_test, sp = TRUE)
k_test <- extract(x = k_okpred, y = k_test, sp = TRUE)
k_test <- extract(x = k_mlrkpred, y = k_test, sp = TRUE)
```

### calculate predicted errors

```{r}
k_test$PE_svm <- k_test$ksvm_pred - k_test$ExK
k_test$PE_mlr <- k_test$kmlr_pred - k_test$ExK
k_test$PE_ok <- k_test$kok_pred - k_test$ExK
k_test$PE_mlrk <- k_test$kmlrk_pred - k_test$ExK
summary(k_test)
```

### calculate RMSE

```{r}
RMSE_svm <- round(sqrt(sum(k_test$PE_svm^2, na.rm=TRUE) / length(k_test$PE_svm)),2)
RMSE_mlr <- round(sqrt(sum(k_test$PE_mlr^2, na.rm=TRUE) / length(k_test$PE_mlr)),2)
RMSE_ok <- round(sqrt(sum(k_test$PE_ok^2, na.rm=TRUE) / length(k_test$PE_ok)), 2)
RMSE_mlrk <- round(sqrt(sum(k_test$PE_mlrk^2, na.rm=TRUE) / length(k_test$PE_mlrk)),2)

paste0("svm: ", RMSE_svm)
paste0("mlr: ", RMSE_mlr)
paste0("mlrk: ", RMSE_mlrk)
paste0("ok: ", RMSE_ok)
```
