## Food Data Prep

### Set project directory, select food data and import

```{r, setup, include=FALSE}
proj_dir <- 'D:/r/KenyaDSM'
knitr::opts_knit$set(root.dir = paste(proj_dir,'data/foods', sep="/"))
# install.packages('tidyverse', 'fs', 'readxl')
```

```{r echo=TRUE}
library(tidyverse)
library(fs)
library(readxl)
library(dplyr)
raw_foods <- as.data.frame(read_excel("food_composition_tables.xlsx", 4))
raw_foods
```

### Clean data

#### 1. Select vegetables and nutrients

```{r}
raw_foods <- raw_foods[307:470, c(2, 12,14,16,17)]
colnames(raw_foods) <- c('Vegetable', 'Calcium', 'Magnesium', 'Potassium', 'Sodium')
raw_foods
```

#### 2. Check and remove nulls

```{r}
print(is.null(raw_foods))
raw_foods <- raw_foods[!(is.na(raw_foods$Vegetable) | raw_foods$Vegetable==""), ]
raw_foods
```

#### 3. Check and remove duplicate veg

```{r}
print(duplicated(raw_foods$Vegetable))
raw_foods <- raw_foods[!duplicated(raw_foods$Vegetable), ] # remove duplicate foods
raw_foods
```

#### 4. Check data types & convert nutrients to floats/ints

```{r}
str(raw_foods)
raw_foods <- raw_foods %>% mutate_at(c('Calcium', 'Magnesium', 'Potassium', 'Sodium'), as.numeric)
str(raw_foods)
```

### EDA

```{r}
summary(raw_foods)
```

```{r}
boxplot(raw_foods[,2:5])
```

Select **K** cause of uniform distribution, reject sodium; compare Mg and Ca since they are similar

```{r}
hist(raw_foods$Magnesium)
hist(raw_foods$Calcium)
```

select **Magnesium** for slightly more varied distribution between Mg and Ca

#### Export veg, potassium and magnesium data

```{r}
write.csv(raw_foods[,c(1,3,4)], "veg_k_mg.csv", row.names=FALSE)
```
