## Food Data Prep

### install packages 
```{r}
install.packages('tidyverse', 'fs', 'readxl')
```

### Set project directory, select food data and import

```{r echo=TRUE}
library(tidyverse)
library(fs)
library(readxl)
library(dplyr)
food_dir <- 'D:/r/KenyaDSM/data/foods/'
raw_foods <- as.data.frame(read_excel(paste(food_dir,"food_composition_tables.xlsx", sep="/"), 4))
raw_foods
```

### Clean data

#### 1. Select vegetables and nutrients

```{r}
raw_foods <- raw_foods[307:470, c(1, 2,12,14,16,17)]
colnames(raw_foods) <- c('Code','Vegetable', 'Calcium', 'Magnesium', 'Potassium', 'Sodium')
raw_foods
```

#### 2. Check and remove nulls
```{r}
sum(is.na(raw_foods$Vegetable) | (raw_foods$Vegetable==""))
raw_foods <- raw_foods[!(is.na(raw_foods$Vegetable) | raw_foods$Vegetable==""), ]
raw_foods
```

#### 3. Check and remove duplicate veg
```{r}
sum(duplicated(raw_foods$Vegetable))
raw_foods <- raw_foods[!duplicated(raw_foods$Vegetable), ] # remove duplicate foods
raw_foods
```

#### 4. Check data types & convert nutrients to floats/ints

```{r}
str(raw_foods)
raw_foods <- raw_foods %>% mutate_at(c('Calcium', 'Magnesium', 'Potassium', 'Sodium'), as.numeric)
str(raw_foods)
```
#### 5. Check and remove canned veg
```{r}
sum(str_detect(raw_foods$Vegetable, 'canned'))
raw_foods <- raw_foods[!str_detect(raw_foods$Vegetable, 'canned'), ] # remove canned foods
raw_foods
```


### EDA
```{r}
summary(raw_foods)
```

```{r}
boxplot(raw_foods[,3:6])
```

Select **K** cause of uniform distribution, reject sodium; compare Mg and Ca since they are similar

```{r}
hist(raw_foods$Magnesium)
hist(raw_foods$Calcium)
```

select **Magnesium** for slightly more varied distribution between Mg and Ca

#### Export veg, potassium and magnesium data
```{r}
write.csv(raw_foods[,c(1,2,4,5)], paste(food_dir,"veg_k_mg.csv",sep="/"), row.names=FALSE)
```