## Soil Data Prep

### Region of Study

-   Identified region: Kenya Method of selection - number of subcounties per county, criterion: closely related for spatial continuity; related to agriculture. Points should also be sparse for variability to cover large extent

-   [Link to Data Source (Kenya Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-ken))

-   Tool : QGIS

-   Selected shapefiles:

    -   counties - ken_admbnda_adm1_iebc_20191031.shp;

### Soil Data

-   Source: Africa Soil Profiles Database

-   Steps in Preparation

    -   Select only Exchangeable Nutrients from Profiles

    -   Select profiles in KE from Layers and Profiles. Select Layer Num 3 - depth of 15cm.

    -   Remove NULL Lat/Long

    -   Do Join of Layers and Profiles by Profile ID to match location and soil nutrients

    -   Select Counties with most profiles and match criteria for region of study : run points in polygon -

```         
QGIS version: 3.34.3-Prizren
QGIS code revision: 47373234ac
Qt version: 5.15.3
Python version: 3.9.18
GDAL version: 3.8.3
GEOS version: 3.12.1-CAPI-1.18.1
PROJ version: Rel. 9.3.1, December 1st, 2023
PDAL version: 2.6.0 (git-version: 3fced5)
Algorithm started at: 2024-03-18T22:06:33
Algorithm 'Count points in polygon' starting…
Input parameters:
{ 'CLASSFIELD' : '', 'FIELD' : 'num_sp', 'OUTPUT' : 'D:/r/KenyaDSM/data/soil/ke_sp/pts_per_cty.shp', 'POINTS' : 'delimitedtext://file:///D:/r/KenyaDSM/data/soil/ke_sp/ke_sp.csv?type=csv&maxFields=10000&detectTypes=yes&xField=loc_longitude&yField=loc_latitude&crs=EPSG:4326&spatialIndex=no&subsetIndex=no&watchFile=no', 'POLYGONS' : 'D:/r/KenyaDSM/data/ke/ken_admbnda_adm1_iebc_20191031.shp', 'WEIGHT' : '' }

No spatial index exists for points layer, performance will be severely degraded
Execution completed in 1.21 seconds
Results:
{'OUTPUT': 'D:/r/KenyaDSM/data/soil/ke_sp/pts_per_cty.shp'}

Loading resulting layers
Algorithm 'Count points in polygon' finished
```

-   Selected counties: Embu and Tharaka Nithi

Run Extract by location from Soil Profiles

```         
QGIS version: 3.34.3-Prizren
QGIS code revision: 47373234ac
Qt version: 5.15.3
Python version: 3.9.18
GDAL version: 3.8.3
GEOS version: 3.12.1-CAPI-1.18.1
PROJ version: Rel. 9.3.1, December 1st, 2023
PDAL version: 2.6.0 (git-version: 3fced5)
Algorithm started at: 2024-03-18T22:26:14
Algorithm 'Extract by location' starting…
Input parameters:
{ 'INPUT' : 'delimitedtext://file:///D:/r/KenyaDSM/data/soil/ke_sp/ke_sp.csv?type=csv&maxFields=10000&detectTypes=yes&xField=loc_longitude&yField=loc_latitude&crs=EPSG:4326&spatialIndex=no&subsetIndex=no&watchFile=no', 'INTERSECT' : 'D:\\r\\KenyaDSM\\data\\soil\\ke_sp\\soil_counties.shp', 'OUTPUT' : 'D:/r/KenyaDSM/data/soil/ke_sp/sample_sp.shp', 'PREDICATE' : [0,6] }

No spatial index exists for input layer, performance will be severely degraded
Execution completed in 0.50 seconds
Results:
{'OUTPUT': 'D:/r/KenyaDSM/data/soil/ke_sp/sample_sp.shp'}

Loading resulting layers
Algorithm 'Extract by location' finished
```

-   Retain ExK and ExMg in soil profiles

#### Covs (GSOC Technical Report/ [FAQ SoilGrids 2017 \| ISRIC](https://www.isric.org/explore/soilgrids/faq-soilgrids-2017))

| Property | Code    | Name                            |
|----------|---------|---------------------------------|
| Climate  | B07CHE3 | Temp Annual Range               |
| Organism | BARL10  | Bareground                      |
| Relief   | DEMENV5 | Land surface elevation          |
| Organism | MAXENV3 | EVI - enhanced vegetation index |
| Climate  | PRSCHE3 | Total annual precipitation      |
| Relief   | SLPMRG5 | Terrain slope                   |
| Relief   | TPIMRG5 | topographic position index      |
| Organism | TREL10  | tree cover                      |
| Climate  | TWIMRG5 | SAGA Wetness Index              |
| Relief   | VDPMRG5 | valley depth                    |
| Climate  | B13CHE3 | ppt of wettest month            |
| Climate  | B14CHE3 | ppt of driest month             |
| Relief   | EXTGSW7 | surface water extent            |
| Climate  | B04CHE3 | Temperature Seasonality         |

-   Clip by extent

```         
QGIS version: 3.34.3-Prizren
QGIS code revision: 47373234ac
Qt version: 5.15.3
Python version: 3.9.18
GDAL version: 3.8.3
GEOS version: 3.12.1-CAPI-1.18.1
PROJ version: Rel. 9.3.1, December 1st, 2023
PDAL version: 2.6.0 (git-version: 3fced5)
Algorithm started at: 2024-03-18T23:49:39
Algorithm 'Clip raster by extent' starting…
Input parameters:
{ 'DATA_TYPE' : 0, 'EXTRA' : '', 'INPUT' : 'D:/r/KenyaDSM/data/covs/full/B14CHE3.tif', 'NODATA' : None, 'OPTIONS' : '', 'OUTPUT' : 'D:/r/KenyaDSM/data/covs/local/driest_ppt.tif', 'OVERCRS' : False, 'PROJWIN' : '37.266925812,38.309360504,-0.916017056,0.068584019 [EPSG:4326]' }

GDAL command:
gdal_translate -projwin 37.266925812 0.068584019 38.309360504 -0.916017056 -of GTiff D:/r/KenyaDSM/data/covs/full/B14CHE3.tif D:/r/KenyaDSM/data/covs/local/driest_ppt.tif
GDAL command output:
Input file size is 960, 1121 
0...10...20...30...40...50...60...70...80...90...100 - done.
Process completed successfully
Execution completed in 0.16 seconds
Results:
{'OUTPUT': 'D:/r/KenyaDSM/data/covs/local/driest_ppt.tif'}

Loading resulting layers
Algorithm 'Clip raster by extent' finished
```

-   Extract extent from covs to add more soil points and enhance prediction

```         
QGIS version: 3.34.3-Prizren
QGIS code revision: 47373234ac
Qt version: 5.15.3
Python version: 3.9.18
GDAL version: 3.8.3
GEOS version: 3.12.1-CAPI-1.18.1
PROJ version: Rel. 9.3.1, December 1st, 2023
PDAL version: 2.6.0 (git-version: 3fced5)
Algorithm started at: 2024-03-19T10:12:20
Algorithm 'Extract layer extent' starting…
Input parameters:
{ 'INPUT' : 'D:/r/KenyaDSM/data/covs/local/annual_ppt.tif', 'OUTPUT' : 'D:/r/KenyaDSM/data/soil/ke_sp/sregion.shp', 'ROUND_TO' : 0 }

Execution completed in 0.77 seconds
Results:
{'OUTPUT': 'D:/r/KenyaDSM/data/soil/ke_sp/soil_region.shp'}

Loading resulting layers
Algorithm 'Extract layer extent' finished

- Drop columns with longitude, latitude in profile; export as CSV with X,Y
```

#### Covariate prep and selection

```{r results = FALSE, message=FALSE, warning=FALSE}
# set directory and epsg
proj_dir <- 'D:/r/KenyaDSM/data'
epsg <- "epsg:4326"

# covariate preparation and selection
library(raster)
library(terra)
library(sf)
library(corrplot)
library(psych)
library(expss)
covs_path <- paste(proj_dir,'covs/local', sep="/")
```

##### create raster stack

```{r}
files <- list.files(path = covs_path, pattern = "tif*$", full.names = TRUE)
covs <- stack(files)
summary(covs) # it is normal to have NoData rasters check Fan et al. use location skipping
```

##### load soil data

```{r}
sp_path <-  paste(proj_dir,'soil/ke_sp', sep="/")
sp_dat <- read.csv(paste(sp_path,'sample/region_sp.csv', sep="/"))
coordinates(sp_dat) <- ~ X + Y
proj4string(sp_dat) <- CRS(epsg)

# confirm that points lie within covariates
{
  plot(covs$sfc_water_extent)
  points(sp_dat)
}
```

##### extract points from covs

```{r}
sp_covs <- as.data.frame(extract(x = covs, y = sp_dat, sp=TRUE))
summary(sp_covs)
```

#### ESDA

-   9999.99s and 0s detected in nutrients (NAs) clean by impute by mean (Baker et al.)

```{r}
sp_covs$ExMg <- na_if(sp_covs$ExMg, -9999.00) # replace -9999 with na
sp_covs$ExK <- na_if(sp_covs$ExK, -9999.00)
sp_covs$ExMg <- na_if(sp_covs$ExMg, 0.00) # replace 0 with na
sp_covs$ExK <- na_if(sp_covs$ExK, 0.00)
# replace per column
sp_covs$ExK[is.na(sp_covs$ExK)] <- mean(sp_covs$ExK, na.rm = T)
sp_covs$ExMg[is.na(sp_covs$ExMg)] <- mean(sp_covs$ExMg, na.rm = T)
summary(sp_covs$ExMg)
summary(sp_covs$ExK)
```
-   drop null columns and rows (location skipping)
```{r}
sp_covs <- subset(sp_covs, select = -c(sfc_water_extent)) 
sp_covs <- sp_covs[complete.cases(sp_covs),]
summary(sp_covs)
```

-   split k and mg each with covs and export

```{r}
k_covs <- sp_covs[-c(4)]
mg_covs <- sp_covs[-c(5)]
write.csv(sp_covs, paste(sp_path,'covs/sp_fullcovs.csv', sep="/"), row.names=FALSE)
write.csv(k_covs, paste(sp_path,'covs/k_fullcovs.csv', sep="/"), row.names=FALSE)
write.csv(mg_covs, paste(sp_path,'covs/mg_fullcovs.csv', sep="/"), row.names=FALSE)
```

##### distribution tests of nutrients

```{r}
hist(k_covs$ExK)
hist(mg_covs$ExMg)
```

-   implies soil has trace amounts of exchangable K compared to exchangable Mg

##### do correlation tests

###### skewness tests to see variation from normal distribution

```{r}
skew(k_covs[4:17], na.rm=TRUE)
skew(mg_covs[4:17], na.rm=TRUE)
```

-   small variations for most covs so apply pearson to select

```{r}
corrplot(cor(k_covs[4:17], method="pearson"), method="color") # do correlation plots
corrplot(cor(mg_covs[4:17], method="pearson"), method="color")
```

-   as matrix format for selection

```{r}
# set to absolute values to capture both +ve and -ve correlations
k_cor <- abs(round(cor(x = as.matrix(k_covs[,4]), y = as.matrix(k_covs[,c(5:17)], method="pearson")),2)) 
mg_cor <- abs(round(cor(x = as.matrix(mg_covs[,4]), y = as.matrix(mg_covs[,c(5:17)], method="pearson")),2))
row.names(k_cor) <- c('k_corr')
row.names(mg_cor) <- c('mg_corr')
k_cor
mg_cor
```
-   select weak to strong correlations: check Mukaka, 2012

```{r}
k_cor <- as.data.frame(k_cor[,k_cor["k_corr",] >= 0.15]) # avoid too weak
mg_cor <- as.data.frame(mg_cor[,mg_cor["mg_corr",] >= 0.2])
k_cor
mg_cor
```

#### create and export modelling sets with selected covariates

```{r}
k_list <- row.names(k_cor) %>% c('X', 'Y', 'ExK') 
mg_list <- row.names(mg_cor) %>% c('X', 'Y', 'ExMg')
k_dat <- subset(k_covs, select=k_list)
mg_dat <- subset(mg_covs, select=mg_list)
write.csv(k_dat, paste(sp_path,"sample/k_model.csv",sep="/"), row.names=FALSE)
write.csv(mg_dat, paste(sp_path,"sample/mg_model.csv",sep="/"),row.names=FALSE)
```
